---
title: "P8105 Homework 2: Data Wrangling"
author: "Catherine Lucey (UNI: cml2270)"
date: 10/10/21
output: html_document
---
## Setup
Only done once, the setup code loads the necessary libraries and double checks that the working directory is correct.

```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
getwd()
```
# Problem 1: Mr. Trash Wheel

## Importing and Cleaning the Mr. Trash Wheel Dataset

```{r import and clean trash man data}
mrtrash_df = read_xlsx("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
           sheet = "Mr. Trash Wheel",
           range = "A2:N408",
           col_names = TRUE
           ) %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, 0))
```

## Importing and Cleaning Precipitaiotn Data for 2018 and 2019

```{r import and read precipitation data}
rain_2018 = read_xlsx("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
          sheet = "2018 Precipitation",
          range = "A2:B14",
          col_names = TRUE
          ) %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  rename(total_precip = total)

rain_2019 = read_xlsx("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
          sheet = "2019 Precipitation",
          range = "A2:B14",
          col_names = TRUE
          ) %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = 2019) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  rename(total_precip = total)

annualrain_df = bind_rows(rain_2018, rain_2019)
```

## Describing Mr. Trash Wheel and Precipitation Data 

The Mr. Trash Wheel dataset provides information on how much trash the Mr. Trash Wheel trash collector picked up from `r min(pull(mrtrash_df, "date"))` to `r max(pull(mrtrash_df, "date"))` in the Baltimore Harbor. The Mr. Trash Wheel dataset gives observations for `r colnames(mrtrash_df)`. In this way, trash collection data are given in total weight (tons) and total volume (cubic yards), as well as weight collected per trash type. These data are presented alongside monthly total inches of rainfall data, which is important because trash is more likely to overflow into the Baltimore harbor after heavy rain. Precipitation data are provided for all `r count(rain_2018)` months in 2018, in which the average total monthly rainfall was `r mean(pull(rain_2018, total_precip))` inches, and `r count(rain_2019)` months in 2019, in which the average total monthly rainfall was `r mean(pull(rain_2019, total_precip))` inches. The cumulative total rainfall in 2019 was `r sum(pull(rain_2019, total_precip))`, and that in 2018 was `r sum(pull(rain_2018, total_precip))`.

On Mr. Trash Wheel's biggest trash pickup day, he collected `r max(pull(mrtrash_df, "weight_tons"))` tons of trash. On Mr. Trash Wheel's least productive day, he only collected `r min(pull(mrtrash_df, "weight_tons"))` tons of trash. The median tons of sports balls Mr. Trash Wheel collected in 2019 was `r median(pull(mrtrash_df, "sports_balls"))`. In collecting all this trash, Mr. Trash Wheel is not only cleaning up the harbor, but also using that waste to generate electricity! On Mr. Trash Wheel's best day, he powered `r max(pull(mrtrash_df, "homes_powered"))` houses.

# Problem 2: FiveThirtyEight

```{r import pols_month and snp data}
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = 
          ifelse(prez_dem == 1, "dem","gop")) %>% 
  select(-prez_dem, -prez_gop, -day) %>% 
  mutate(year = as.numeric(year))

snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  mutate(date = mdy(date)) %>%
  separate(date, into = c("year","month", "day")) %>%
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(68 > year %% 100 & 16 > year %% 100, 2000 + year %% 100, 1900 + year %% 100)) %>%
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-day) %>% 
  relocate(year)
```

```{r tidying unemployment data}

unemploy_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate") %>% 
  mutate(month = recode(month, jan = "January", feb = "February", mar = "March", apr = "April", may = "May", jun = "June", jul = "July", aug = "August", sep = "September", oct = "October", nov = "November", dec = "December" ))
```

```{r joining all three datasets}

unemp_pols_snp_join = inner_join(snp_df, pols_df, by = c("year", "month")) %>% 
  inner_join(unemploy_df, by = c("year", "month"))

```

## Describing the FiveThirtyEight Data

The three Five Thirty Eight datasets show the change over time in politics, as measured by the balance of elected officials in the democrat and republican parties and the strength of the economy, via the closing value of the S&P 500 index and the unemployment rate. The politics dataset has `r nrow(pols_df)` rows and `r ncol(pols_df)` columns, the unemployment dataset has `r nrow(unemploy_df)` rows and `r ncol(unemploy_df)` columns, and the S&P 500 dataset has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns.

The unemployment dataset gives unemployment rates from `r min(pull(unemploy_df, "year"))` to `r max(pull(unemploy_df, "year"))`, the S&P 500 dataset gives the closing value of the stock index from `r min(pull(snp_df, "year"))` to `r max(pull(snp_df, "year"))`, and lastly the politics dataset gives how many senators, governors, and representatives were democrats vs republicans, as well as whether the president was republican or democrat, from `r min(pull(pols_df, "year"))` to `r max(pull(pols_df, "year"))`. 


# Problem 3: Baby Names

```{r import and tidy babynames}

babyname_df = read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct(.keep_all = TRUE) %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>% 
  mutate(gender = str_to_lower(gender)) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>% 
  mutate(ethnicity = recode(ethnicity, `asian and paci` = "asian and pacific islander", `black non hisp` = "black non hispanic", `white non hisp` = "white non hispanic"))
  
```

```{r olivia plot}

olivia_df = babyname_df %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-gender) %>%  
  pivot_wider(names_from = year_of_birth, values_from = count)
  
knitr::kable(olivia_df)
print(olivia_df)

```

```{r most popular boy names}

boys_df = babyname_df %>% 
  filter(gender == "male", rank == "1") %>% 
  pivot_wider(names_from = year_of_birth, values_from = childs_first_name) %>% 
  select(-gender)

```

```{r}
boys_subset_df = babyname_df %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016")
ggplot(boys_subset_df, aes(x = rank, y = count)) + geom_point() + ylab("Number of Children With Name") + xlab("Popularity Rank of Name")
```

